### 2020-07-27 Catuscia - final/07-discussion.tex
da dropbox

Our encoding increases the expressivity of  RS  concerning:
%the behaviour of the context,
the possibility of alternative behaviour of
mutated entities, and  the communication between two different Reaction Systems.
It is important to note that our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one transition comes out, 
as the dynamics is totally deterministic.

%\subsection{Recursive contexts}
%In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
% which are the entities that are provided from the context. 
%Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
%Then, the context behaviour would not necessarily end after $n$ steps,
%and  could be infinite.
%For example, in an extended interactive process, we may want that the entity $s$ is 
%intermittently provided by the context every two steps: 
%\[
%\begin{array}{lcll}
%Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
%Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
%Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
%\end{array}
%\]

\subsection{Mutating entities} 
In RS, when an entity is present, it can potentially be involved in each reactions where it is required.
With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
an entity, in a way that the mutated version of the entity can take 
part to only a subset of the rules requiring  the  \emph{normal version} of the entity.
For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
Reaction $a1$ produces also $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
say $s1'$.
When $s1'$ is produced, reaction $a2$ behaves in the same way as if $s1$ would be absent, whereas $a2$
recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
Technically, in both cases it is enough to add one more non 
deterministic choice in the code of $P_{a1}$ and $P_{a2}$.

\subsection{Communicating Reaction Systems}
We sketch how it is possible to program two RSs encodings, in a way that
the entities that usually come from context of one RS will be provided instead from the other RS.

@@

\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline
$rs1$  & $rs2$ \\
\hline
%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
$a_1=(s,\ , x)$ & $a_2=(y,\  ,s)$\\
%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
\hline 
\end{tabular}
\caption{The two Reaction Systems $rs1$ and $rs2$.}
\label{tab:rules}
\end{table}
 \noindent
Now, we set our example such that the two contexts, for  $rs1$ and $rs2$, do not provide any entities.
We also assume that entity $s$ in $rs1$ is provided by $rs2$, as $rs2$ produces a quantity of $s$ that is enough for $rs1$ and $rs2$.
For technical reasons, we can not use the same name for  $s$ in both the two RSs, then we use the name $ss$ in $rs2$.
%
We need to modify our translation technique to suite this new setting. 
As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the absence of entities. Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
%Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
%Please also note that, for our simple example, we do not have anymore to specify the all steps
%of a interactive system, as the communication mechanism from $rs2$ to $rs1$ will generate an infinite
%behaviour.
%\noindent
%First, we translate the rules in $rs1$:
First, we translate the reaction in $rs1$:
\[
\begin{array}{lcl}
%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\tilde{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{r1_2}.P_{r1_1} +
%\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
\llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\widetilde{x}_o}\chainend{a_2}.P_{a_1} +
\startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
%\end{array}
%\]
%\\
%
%\begin{array}{lcl}
%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\tilde{s1}}{\noact}\chainedlink{\noact}{\tilde{s1}}\chainend{r2_1}.P_{r2_1} +
%\startchain{r1_2}\chainedlink{s2}{\noact}\chainedlink{\noact}{s2}\chainend{r2_1}.P_{r2_1}\\
\end{array}
\]

\noindent
Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
The translation for the entities follows.
\[
\begin{array}{lclcl}
\llbracket s \rrbracket &\defeq &P_{s} &\defeq & \startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\widehat{s}_i}}\chainend{{\color{blue}\widehat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}\\
&& \overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\widehat{s}_i}\chainend{\widehat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}\\
%\begin{array}{lclclll}
\llbracket x \rrbracket &\defeq & P_{x}& \defeq &\startchain{\widetilde{x}_i}\chainend{\widetilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\\
%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +
%\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
%\\
%\\
&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\widetilde{x}_i}}\chainend{{\color{gray}\widetilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\end{array}
%\]
%
%\[

@@

\[
\begin{array}{lcl}
\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\widetilde{ss}_o}\chainend{\silent}.P_{a_2}+
\startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
\chainedlink{dss_i}{\noact}\chainedlink{\noact}{dss_o}\chainend{\silent}.P_{a_2}\\
%\end{array}
%\]
%
%\[
%\begin{array}{lcl}
%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\tilde{s3}}{\noact}\chainedlink{\noact}{\tilde{s3}}\chainend{\silent}.P_{r2_2}
%+
%\startchain{r2_2}\chainedlink{\overline{ss1}}{\noact}\chainedlink{\noact}{\overline{ss1}}\chainend{\silent}.P_{r2_2}\\
\end{array}
\]
\noindent
In the translation of the entities in  $rs2$, we introduce the mechanism that allows the entity
$s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
created to synchronise with $rs1$   on link $\link{\widehat{s}_i}{\widehat{s}_o}$:
\[
\begin{array}{lclcl}
\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\widetilde{ss}_i}\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\chainend{\widetilde{ss}_o}.P_{ss}+
\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}} \\

%\end{array}
%\]
%\[
%\begin{array}{lcllcl}
&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\widetilde{ss}_i}}\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\widehat{s}_o}}\chainend{{\color{red}\widetilde{ss}_o}}.P_{ss}
+
\startchain{dss_i}\chainend{dss_o}.\overline{P_{ss}}\\
\llbracket y\rrbracket&\defeq & P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
&&\overline{P_{y}}&\defeq &
%\startchain{\overline{y}_i}\chainedlink{\overline{y}_o}{\noact}\chainedlink{\noact}{\tilde{y}_i}\chainend{\tilde{y}_o}.P_{y}
%+ 
\startchain{\overline{y}_i}\chainend{\overline{y}_o}.\overline{P_{y}}
\end{array}
\]
%\[
%\begin{array}{lcllcl}

@@

\noindent
We now assume that the initial system is $S \defeq \restrict{names}(P_{a_1}|P_{a_2}|P_s|P_y|\overline{P_x}|\overline{P_{ss}})$, i.e.
only entities $s$ and $y$ are present. Now, the only possible transition has  the following label (that we report without restriction):
\[
\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\widetilde{x}_i}{{\color{gray}\widetilde{x}_i}}\chainedlink{{\color{gray}\widetilde{x}_o}}{\widetilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\widetilde{ss}_i}{{\color{red}\widetilde{ss}_i}}
\chainedlink{{\color{red}\widehat{s}_i}}{{\color{blue}\widehat{s}_i}}
\chainedlink{{\color{blue}\widehat{s}_o}}{{\color{red}\widehat{s}_o}}
\chainedlink{{\color{red}\widetilde{ss}_o}}{\widetilde{ss}_o}\chainend{\silent},
\]
where the black links belong to the prefixes of $P_{a_1}$, and $P_{a_2}$, the blue links belong to $P_s$, the gray links belong to $P_y$, and $\overline{P_x}$ and the red links belong to $\overline{P_{ss}}$.
After the execution, the entity $s$ is still present in $rs1$ as it has been provided by $rs2$.\\

As we have briefly sketched, our model of two 
 \emph{communicating Reaction Systems} can enable the study of the behaviour of one RS in relation to another one. 
Thus, the products of the reactions of one RS can become the input for another one. 
This could allow for a modular approach to modeling complex systems, by composing different Reaction Systems.

%which can be modified to represent pathological states as well.

% Again, we stress the fact the verbosity of the {\tt link}-calculus code 
%should not be
% seen as a drawback, as it undergoes to automatic code translation.

>>>>

Our encoding increases the expressivity of  RS  concerning:
the behaviour of the context, the possibility of alternative behaviour of
mutated entities and the communication between two different reaction systems.
It is important to note that our encoding guarantees that from each state, 
in the  c\CNA \ transition system, only one transition comes out, 
as the dynamics is totally deterministic.

\subsection{Recursive contexts}
In RS, the behaviour of the context is finite. For the first $n$ steps, it is specified 
 which are the entities that are provided from the context. 
Using c\CNA\ we can describe in a natural way the behaviour of the context in a recursive way. 
Then, the context behaviour would not necessarily end after $n$ steps,
and  could be infinite.
For example, in an extended interactive process, we may want that the entity $s$ is 
intermittently provided by the context every two steps: 
\[
\begin{array}{lcll}
Cxt_s & \defeq &  \startchain{cxt_j}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{cxt_{j+1}}.Cxt_{s'}, &\mbox{ context provides $s$;} \\
Cxt_{s'} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s''},& \mbox{ context doesn't provide $s$;} \\
Cxt_{s''} & \defeq & \startchain{cxt_j}\chainedlink{\underline{s}_i}{\noact}\chainedlink{\noact}{\underline{s}_o}\chainend{cxt_{j+1}}.Cxt_{s}, & \mbox{ context doesn't provide $s$.} \\
\end{array}
\]

\subsection{Mutating entities} 
In RS, when an entity is present, it can potentially be involved in each reaction where it is required.
With a few more lines of code, in c\CNA~it is possible to describe the behaviour of a mutation of 
an entity, in a way that the mutated version of the entity can take 
part to only a subset of the rules requiring  the  \emph{normal version} of the entity.
For example, let us assume that entity $s1$ is consumed by reactions $a1$ and $a2$.
Reaction $a1$ produces $s1$ if $s2$ is present, otherwise $a1$ produces a mutated version of $s1$, 
say $s1'$.
When $s1'$ is produced, reaction $a1$ behaves in the same way as if $s1$ would be absent, whereas $a2$
recognises the presence of $s1'$ and behaves in the same way as if $s1$ would be present.
Technically, in both cases it is enough to add one more non 
deterministic choice in the code of $P_{a1}$ and $P_{a2}$.

\subsection{Communicating reaction systems} 
We sketch how it is possible to program two RS encodings, in a way that 
the entities that usually come from context of one RS will be provided instead from the other RS.

@@

\begin{table}[t]
\centering
\begin{tabular}{|c|c|}
\hline
$rs1$  & $rs2$ \\
\hline
%$r1_1=(s1,\ , s2)$ & $r2_1=(s3,\  ,s1)$\\
$a_1=(s,\ , x)$ & $a_2=(y,\  ,s)$\\
%$r1_2=(\ , s2,s1)$ & $r2_2=(s1,\ ,s3)$\\
\hline 
\end{tabular}
\caption{The two reaction systems $rs1$ and $rs2$.}
\label{tab:rules}
\end{table}
 \noindent
Now, we set our example such that the two contexts, for  $rs1$ and $rs2$, do not provide any entities.
We also assume that entity $s$ in $rs1$ is provided by $rs2$, as $rs2$ produces a quantity of $s$ that is enough for $rs1$ and $rs2$.
For technical reasons, we can not use the same name for  $s$ in both the two RSs, then we use the name $ss$ in $rs2$.
%
We need to modify our translation technique to suite this new setting. 
As we do not model contexts, we introduce  \emph{dummy} channel names $dx$ and $dss$ to model the case $x$ and $ss$ are not produced.
%absence of entities. 
Also, thanks to the simplicity of the example, we can leave out the use of the   $p_i$ channels. This streamlining does not affect the programming technique we propose to make two RSs communicate.
%Thus, there will be no $Cxt$ processes and we manage the code in a way that $rs1$ will receive the entity $s1$ from $rs2$. 
%Please also note that, for our simple example, we do not have anymore to specify the all steps
%of a interactive system, as the communication mechanism from $rs2$ to $rs1$ will generate an infinite
%behaviour.
%\noindent
%First, we translate the rules in $rs1$:
First, we translate the reaction in $rs1$:
\[
\begin{array}{lcl}
%\llbracket r1_1\rrbracket \defeq P_{r1_1}&\defeq &\startchain{\silent}\chainedlink{s1}{\noact}\chainedlink{\noact}{s1}\chainedlink{\tilde{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{r1_2}.P_{r1_1} +
%\startchain{\silent}\chainedlink{\overline{s1}}{\noact}\chainedlink{\noact}{\overline{s1}}\chainend{r1_2}.P_{r1_1}\\
\llbracket a_1\rrbracket \defeq P_{a_1}&\defeq &
\startchain{\silent}\chainedlink{s_i}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{\tilde{x}_o}\chainend{a_2}.P_{a_1} +
\startchain{\silent}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}    \chainedlink{dx_i}{\noact}\chainedlink{\noact}{dx_o}\chainend{a_2}.P_{a_1}\\
%\end{array}
%\]
%\\
%
%\begin{array}{lcl}
%\llbracket r1_2\rrbracket \defeq P_{r1_2}&\defeq &\startchain{r1_2}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\overline{s2}}\chainedlink{\tilde{s1}}{\noact}\chainedlink{\noact}{\tilde{s1}}\chainend{r2_1}.P_{r2_1} +
%\startchain{r1_2}\chainedlink{s2}{\noact}\chainedlink{\noact}{s2}\chainend{r2_1}.P_{r2_1}\\
\end{array}
\]

\noindent
Please note, that prefixes of process $P_{a_1}$ end with the channel name $a_2$, as the link chain is now connected with the  reaction of  $rs2$.
The translation for the entities follows.
\[
\begin{array}{lclcl@{\qquad}lclcl}
\llbracket s \rrbracket &\defeq &P_{s}&\defeq &\startchain{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{\noact}\chainedlink{\noact}{{\color{blue}\hat{s}_i}}\chainend{{\color{blue}\hat{s}_o}}.P_{s} + \startchain{s_i}\chainend{s_o}.\overline{P_{s}}
&
%\begin{array}{lclclll}
\llbracket x \rrbracket &\defeq& P_{x}&\defeq &\startchain{\tilde{x}_i}\chainend{\tilde{x}_o}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\\
%&&\overline{P_{s2}}&\defeq &\startchain{\overline{s2}}\chainedlink{\overline{s2}}{\noact}\chainedlink{\noact}{\tilde{s2}}\chainend{\tilde{s2}}.P_{s2} +
%\startchain{\overline{s2}}\chainend{\overline{s2}}.\overline{P_{s2}}\\
%\\
%\\
&&\overline{P_{s}}&\defeq &\startchain{\overline{s}_i}\chainedlink{\overline{s}_o}{\noact}\chainedlink{\noact}{\hat{s}_i}\chainend{\hat{s}_o}.P_{s}+\startchain{\overline{s}_i}\chainend{\overline{s}_i}.\overline{P_{s}}
&
&& \overline{P_{x}}&\defeq &\startchain{{\color{gray}\tilde{x}_i}}\chainend{{\color{gray}\tilde{x}_o}}.P_{x} +  \startchain{dx_i}\chainend{dx_o}.\overline{P_{x}}\\
%\end{array}
%\]
%
%\[

@@

\[
\begin{array}{lcl}
\llbracket a_2\rrbracket \defeq P_{a_2}&\defeq &\startchain{a_2}\chainedlink{y_i}{{\color{gray}\noact}}\chainedlink{{\color{gray}\noact}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\noact}}\chainedlink{{\color{red}\noact}}{\tilde{ss}_o}\chainend{\silent}.P_{a_2}+
\startchain{a_2}\chainedlink{\overline{y}_i}{\noact}\chainedlink{\noact}{\overline{y}_o}    
\chainedlink{dss_i}{\noact}\chainedlink{\noact}{dss_o}\chainend{\silent}.P_{a_2}\\
%\end{array}
%\]
%
%\[
%\begin{array}{lcl}
%\llbracket r2_2\rrbracket \defeq P_{r2_2}&\defeq &\startchain{r2_2}\chainedlink{ss1}{\noact}\chainedlink{\noact}{ss1}\chainedlink{\tilde{s3}}{\noact}\chainedlink{\noact}{\tilde{s3}}\chainend{\silent}.P_{r2_2}
%+
%\startchain{r2_2}\chainedlink{\overline{ss1}}{\noact}\chainedlink{\noact}{\overline{ss1}}\chainend{\silent}.P_{r2_2}\\
\end{array}
\]
\noindent
In the translation of the entities in  $rs2$, we introduce the mechanism that allows the entity
$s$ ($ss$ in $rs2$) to be provided in $rs1$. Every time $ss$ is produced in $rs2$,  a virtual link is
created to synchronise with $rs1$   on link $\link{\hat{s}_i}{\hat{s}_o}$:
\[
\begin{array}{lclcl@{\qquad} lclcl}
\llbracket ss \rrbracket& \defeq& P_{ss}&\defeq &\startchain{\tilde{ss}_i}\chainedlink{\hat{s}_i}{\noact}\chainedlink{\noact}{\hat{s}_o}\chainend{\tilde{ss}_o}.P_{ss}+
\startchain{dss_i}   \chainend{dss_o}.\overline{P_{ss}}
&
\llbracket y\rrbracket&\defeq & P_{y}&\defeq &\startchain{{\color{gray}y_i}}\chainend{{\color{gray}y_o}}.\overline{P_{y}} \\
%\end{array}
%\]
%\[
%\begin{array}{lcllcl}
&&\overline{P_{ss}}&\defeq &\startchain{{\color{red}\tilde{ss}_i}}\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\noact}}\chainedlink{{\color{blue}\noact}}{{\color{red}\hat{s}_o}}\chainend{{\color{red}\tilde{ss}_o}}.P_{ss}
+
\startchain{dss_i}\chainend{dss_o}.\overline{P_{ss}}
&
&&\overline{P_{y}}&\defeq &
%\startchain{\overline{y}_i}\chainedlink{\overline{y}_o}{\noact}\chainedlink{\noact}{\tilde{y}_i}\chainend{\tilde{y}_o}.P_{y}
%+ 
\startchain{\overline{y}_i}\chainend{\overline{y}_o}.\overline{P_{y}}
\end{array}
\]
%\[
%\begin{array}{lcllcl}

@@

\noindent
We now assume that the initial system is $S \defeq \restrict{names}(P_{a_1}|P_{a_2}|P_s|P_y|\overline{P_x}|\overline{P_{ss}})$, i.e.
only entities $s$ and $y$ are present. Now, the only possible transition has  the following label (that we report without restriction):
\[
\startchain{\silent}\chainedlink{s_i}{{\color{blue}s_i}}\chainedlink{{\color{blue}s_o}}{s_o}\chainedlink{\tilde{x}_i}{{\color{gray}\tilde{x}_i}}\chainedlink{{\color{gray}\tilde{x}_o}}{\tilde{x}_o} \chainedlink{a_2}{a_2}\chainedlink{y_i}{ {\color{gray}y_i}}\chainedlink{{\color{gray}y_o}}{y_o}\chainedlink{\tilde{ss}_i}{{\color{red}\tilde{ss}_i}}
\chainedlink{{\color{red}\hat{s}_i}}{{\color{blue}\hat{s}_i}}
\chainedlink{{\color{blue}\hat{s}_o}}{{\color{red}\hat{s}_o}}
\chainedlink{{\color{red}\tilde{ss}_o}}{\tilde{ss}_o}\chainend{\silent},
\]
where the black links belong to the prefixes of $P_{a_1}$, and $P_{a_2}$, the blue links belong to $P_s$, the gray links belong to $P_y$, and $\overline{P_x}$ and the red links belong to $\overline{P_{ss}}$.
After the execution, the entity $s$ is still present in $rs1$ as it has been provided by $rs2$.\\

As we have briefly sketched, our model of two 
\emph{communicating reaction systems} 
can enable the study of the behaviour of one RS
in relation to another one.
% In Section~\ref{ex:lactose} describing  the behaviour of the \emph{lac} operon,
% the two entities lactose and glucose are controlled 
% non deterministically by the context.
% In our framework,
% instead, by exploiting the expressivity of c\CNA, 
% the \emph{lac} operon system can be connected with the two systems 
% producing the lactose and the glucose.
% This way, the presence of these two entities in the \emph{lac} operon system 
% could be regulated by realistic mechanisms.
Thus, the products of the reactions of one RS can become the input for
another one. This could allow for a modular approach to modeling 
complex systems, by composing different Reaction Systems.

### 2020-07-27 Catuscia - submitted/01-introduction.tex
da dropbox

Natural Computing is an emerging area of research which has two main 
aspects: human designed computing inspired by nature, and computation 
performed in nature. Reaction Systems (RSs)~\cite{BEMR11} are a 
% model of
% formal computation 
rewriting formalism
%model 
inspired by the way biochemical reactions take place in living 
cells.  
This theory has already shown to be relevant in several different 
fields, such as computer science~\cite{MPR15}, 
biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, and
molecular chemistry~\cite{OY16}.
%,
% and it
%is continuously developing, with a growing number of workshops and
%international meetings. 
Reaction Systems formalise the mechanisms of biochemical systems, 
such as {\em facilitation} and {\em inhibition}. 
As a qualitative approximation of the real biochemical reactions, they
% do not take
%into account the quantity of reagents effectively present. They
consider if a necessary reagent is or not present, and likewise they
consider if an inhibiting molecule is or not present. 
The  possible reactants and inhibitors are called `entities'.
RSs model in a direct way the interaction of a living cell
with the environment (called `context'). However, two RSs are seen
as independent models and do not interact.

%Formally, Reaction Systems are defined as a rewriting formalism.
In this paper, we present an encoding from RSs, 
to the open multiparty process algebra c\CNA\footnote{After `chained Core Network 
Algebra'.}, a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
This formalism allows several processes to synchronise and 
communicate altogether, at the same time, with a new communicating
mechanism based on links and link chains. 
Our initial motivation for introducing this mechanism was to encode
Mobile Ambients~\cite{CardelliG00}, obtaining a much stronger operational
correspondence than any one available in the literature, such as the one in~\cite{B16}.
 This allowed us to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
Process calculi have been used successfully to model
biological processes, see~\cite{BBDFH18} for a recent survey.
We illustrate our embedding by means of some simple basic examples,
and then we consider a more complex example, by modeling a RS 
representing a regulatory network for \emph{lac} operon, presented in~\cite{CMMBM12}. We also show that our embedding preserves the main
features of RSs, and prove its correctness and completeness.
Our main contributions are as follows:
\begin{itemize}
\item the behaviour of the context, for each single 
entity,
%\footnote{Entities that belong to the `domain' of a RS.}
can be
specified in a recursive way as an ordinary process;
\item 
%\begin{color}{red}
our translation results in  a  c\CNA \ system where from
each state only one transition can be generated, thus the c\CNA \
computation is fully deterministic as that of the encoded RS;
%\end{color}
\item we can express the behaviour of entity mutation,
%change the behaviour of  each single entity $s$  in a
%way that a mutation $s'$ can be expressed; 
in such a way that the mutated entity
$s'$ can  take part to only a subset of rules requiring entity $s$;
\item with a little coding effort, two RSs can
communicate; i.e.  a subset of those entities that the context  can
provide, are then provided by a second RS.
\end{itemize}

@@

\paragraph{Structure of the paper.} Section \ref{sec:rs} 
describes RSs and their semantics (interactive processes).
Section \ref{sec:ccna} describes briefly the c\CNA \ process algebra and
its operational semantics.
Section \ref{sec:trans} defines the embedding of RSs in
c\CNA \ processes and shows some simple examples to illustrate 
it.
Section \ref{ex:lactose} shows a more complex example taken from
the literature on RSs and illustrates a \emph{lac} operon.
Section~\ref{sec:discussion} presents some features and 
advantages of our embedding for the compositionality of RSs.
Finally, Section~\ref{sec:conclusion}
discusses future work and concludes.

>>>>

Natural Computing is an emerging area of research which has two main 
aspects: human designed computing inspired by nature, and computation 
performed in nature. Reaction Systems (RSs)~\cite{BEMR11} are a 
% model of
% formal computation 
rewriting formalism
%model 
inspired by the way biochemical reactions take place in living 
cells.  
This theory has already shown to be relevant in several different 
fields, such as computer science~\cite{MPR15}, 
biology~\cite{ABP14,CMMBM12,Az17,BarbutiGLM16}, 
molecular chemistry~\cite{OY16}.
%,
% and it
%is continuously developing, with a growing number of workshops and
%international meetings. 
Reaction Systems formalise the mechanisms of biochemical systems, 
such as {\em facilitation} and {\em inhibition}. 
As a qualitative approximation of the real biochemical reactions, they
% do not take
%into account the quantity of reagents effectively present. They
consider if a necessary reagent is or not present, and likewise they
consider if an inhibiting molecule is or not present. 
The  possible reactants and inhibitors are called `entities'.
RSs model in a direct way the interaction of a living cell
with the environment (called `context'). However, two RSs are seen
as independent models and do not interact.

%Formally, Reaction Systems are defined as a rewriting formalism.
In this paper, we present an encoding from RSs, 
to the open multiparty process algebra c\CNA,\footnote{After `chained Core Network 
Algebra'.} a variant of the {\tt link}-calculus~\cite{BodeiBB12,BBB17} without name mobility.
This formalism allows several processes to synchronise and 
communicate altogether, at the same time, with a new communicating
mechanism based on links and link chains. 
Our initial motivation for introducing this mechanism was to encode
Mobile Ambients~\cite{CardelliG00}, getting a much stronger operational
correspondence than any available in the literature, such as the one in~\cite{B16}.
 This allowed us to easily encode calculi for biology equipped with membranes, as in~\cite{BodeiBBC14}.
Process calculi have been used successfully to model
biological processes, see~\cite{BBDFH18} for a recent survey.
We illustrate our embedding by means of some simple basic examples,
and then we consider a more complex example, by modeling a RS 
representing a regulatory network for \emph{lac} operon, presented in~\cite{CMMBM12}. We also show that our embedding preserves the main
features of RSs, and prove its correctness and completeness.
Our main contributions are as follows:
\begin{itemize}
\item the behaviour of the context, for each single 
entity,
%\footnote{Entities that belong to the `domain' of a RS.}
can be
specified in a recursive way as an ordinary process;
\item we can express the behaviour of entity mutation,
%change the behaviour of  each single entity $s$  in a
%way that a mutation $s'$ can be expressed; 
in such a way that the mutated entity
$s'$ can  take part to only a subset of rules requiring entity $s$;
\item with a little coding effort, two RSs can
communicate; i.e.  a subset of those entities that the context  can
provide, are then provided by a second RS;
\item as our translation results in  a  c\CNA \ system, from
each state only one transition can be generated, thus the c\CNA \
computation is fully deterministic.
\end{itemize}

@@

\paragraph{Structure of the paper.} Section \ref{sec:rs} 
describes RSs and their semantics (interactive processes).
Section \ref{sec:ccna} describes briefly the c\CNA \ process algebra and
its operational semantics.
Section \ref{sec:trans} defines the embedding of RSs in
c\CNA \ processes and shows some simple examples to illustrate 
it.
Section \ref{ex:lactose} shows a more complex example taken from
the literature on RSs and illustrating a \emph{lac} operon.
Section~\ref{sec:discussion} presents some features and 
advantages of our embedding for the compositionality of RSs.
Finally, Section~\ref{sec:conclusion}
discusses future work, and concludes.
